---
title: "ECC exploratory analysis"
output: html_notebook
author: "Bose Falk"
---


```{r}
library(ggplot2)
library(tidyverse)
source('lib/loading_scripts.R')

dat <- load_data('trust')
validate_data(dat)

```

Analyze only one round for now:
```{r}
dat %<>% filter(essround == "4")

```

As response variables, we're interested in the trust people have in their government institutions. First we explore simply just trust in the country's parliament, variable `trstprl`.

Look at 'time spent consuming news or programmes about politics or current affairs' for different media types in different countries (newspaper, radio, TV):

```{r}

plot_df <- dat %>% select(cntry, trstprl, tvpol, rdpol, nwsppol) %>% gather('media', 'score', tvpol, rdpol, nwsppol)


ggplot(plot_df %>% drop_na, aes(x = as.factor(trstprl), y = score)) + geom_count() + facet_grid(media ~ cntry)


```

Overall, trust is higher in Denmark and Estonia, amount of politics-related programming watched is about the same, and no clear indicator overall that there is an upward trend.

Summary variable for consumption of political-related programming across all three media types, `pol_prog_avg`, simple sum. The question is worded in 'time spend watching...' (not relative), so the sum should give a good indication of the total time.

```{r}

dat %<>% mutate(pol_prog_avg = rowSums(select(., nwsppol, rdpol, tvpol)))

ggplot(dat %>% drop_na, aes(x = trstprl, y = pol_prog_avg)) + 
  geom_count() + 
  geom_smooth(method='lm') +
  facet_grid(cntry ~ .)

fit2 <- dat %>% group_by(cntry) %>% do(mod = lm(trstprl ~ pol_prog_avg, data = .))
fit2$mod

```


### Intra-person difference in trust variables:

```{r}
dat_diff <- dat %>% mutate(difflgl = trstprl - trstlgl, 
                           diffplc = trstprl - trstplc, 
                           diffplt = trstprl - trstplt, 
                           diffprt = trstprl - trstprl,
                           diffep = trstprl - trstep,
                           diffun = trstprl - trstun)
dat_diff_plot <- dat_diff %>% gather('comparison', 'difference', 31:36)
ggplot(dat_diff_plot, aes(x = comparison, y = difference)) + geom_boxplot() + facet_grid(cntry ~ .)



```


```{r}
plot_df <- dat %>% select(cntry, trstprl, trstlgl, trstplc, trstprt) %>%
  rename(country = cntry, 
         trust_legal = trstlgl, 
         trust_polparties = trstprt, 
         trust_police = trstplc, 
         trust_parliament = trstprl)
plot_df <- gather(plot_df, 'variable', 'survey_level', 2:5)

ggplot(plot_df, aes(x = variable, y = survey_level, fill = country)) +
  geom_boxplot() +
  labs(title = "Trust variable survey answers", 
       x = "Survey Question", 
       y = "Response")


```