---
title: "ECC exploratory analysis"
output: html_notebook
author: "Bose Falk"
---


```{r}
library(ggplot2)
library(tidyverse)
source('lib/loading_scripts.R')

dat <- load_data('trust')
validate_data(dat)

```

As response variables, we're interested in the trust people have in their government institutions. First we explore simply just trust in the country's parliament, variable `trstprl`.

Look at 'time spent consuming news or programmes about politics or current affairs' for different media types in different countries (newspaper, radio, TV):

```{r}

plot_df <- dat %>% select(cntry, trstprl, tvpol, rdpol, nwsppol) %>% gather('media', 'score', tvpol, rdpol, nwsppol)


ggplot(plot_df %>% drop_na, aes(x = as.factor(trstprl), y = score)) + geom_count() + facet_grid(media ~ cntry)


```

Overall, trust is higher in Denmark and Estonia, amount of politics-related programming watches is about the same, and no clear indicator overall that there is an upward trend.

Summary variable for consumption of political-related programming across all three media types, `pol_prog_avg`, simple sum. The question is worded in 'time spend watching...' (not relative), so the sum should give a good indication of the total time.

```{r}

dat %<>% mutate(pol_prog_avg = rowSums(select(., nwsppol, rdpol, tvpol)))

ggplot(dat %>% drop_na, aes(x = trstprl, y = pol_prog_avg)) + 
  geom_count() + 
  geom_smooth(method='lm') +
  facet_grid(cntry ~ .)

fit2 <- dat %>% group_by(cntry) %>% do(mod = lm(trstprl ~ pol_prog_avg, data = .))
fit2$mod

```